SDSS-IV MaNGA: The Universal Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-07-17 maintex.tex
title change

\title{SDSS-IV MaNGA: Radial Profiles of Specific Star Formation Rate in Nearby Interacting Galaxies}

@@

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of 160 interacting galaxies within the MaNGA (Mapping Nearby Galaxies at Apache Point Observatory) survey. We find that the specific star formation rate (sSFR) is centrally enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We find that the level of the central enhancement is independent of galaxy mass. Low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. Assuming a constant burst merger induced star formation over a 1 Gyr period, the mass growth rate in the centers of the paired galaxies will be 1.6$-$2.3$\times$ higher than secular mass growth. Since the merger induced sSFR enhancement is constant over stellar mass, low mass galaxies will experience a higher fractional mass growth rate which is $\sim$4.3 higher than that of high mass galaxies. This indicates that merger induced star formation has a significant impact on the mass evolution of low mass galaxies.

>>>>

\title{SDSS-IV MaNGA: The Universal Profile of Enhanced Star Formation in Close Galaxy Pairs}

@@

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of 160 interacting galaxies within the MaNGA (Mapping Nearby Galaxies at Apache Point Observatory) survey. We find that the specific star formation rate (sSFR) is centrally (e.g. within a single effective radius) enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We split the radial profile into four separate stellar mass bins from \logm\ $=$ 9.0$-$11.0 and find that the sSFR enhancement has no dependence on the stellar mass of the paired galaxies. Within one effective radius, the sSFR enhancement in each of the mass bins falls within $\sim$0.1 dex of the median profile for the whole pair sample. Low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. Assuming a constant burst merger induced star formation over a 1 Gyr period, the mass growth rate in the centers of the paired galaxies will be 1.6$-$2.3$\times$ higher than secular mass growth. Since the merger induced sSFR enhancement is constant over stellar mass, low mass galaxies will experience a higher fractional mass growth rate which is $\sim$4.3 higher than that of high mass galaxies. This indicates that merger induced star formation has a significant impact on the mass evolution of low mass galaxies.
